---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview

`maxnodf` calculates the maximum NODF value that can be achieved in a bipartite network with a given number of rows, columns and links.

## Installation

You can install `maxnodf` from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools") # install devtools if needed
devtools::install_github("CHoeppke/maxnodf")
```

## Use
`maxnodf` has one function: `maxnodf()`. For a given network, `maxnodf` calculates the maximum nestedness that can be achieved in a network with a given number of rows, columns and links, subject to the constraint that all rows and columns must have at least one link (i.e. row and column totals must always be >= 1). As input, `maxnodf()` takes either a numeric matrix describing a bipartite network (a bipartite incidence matrix where elements are positive numbers if nodes interact, and 0 otherwise) or a numeric vector of length three of the form `c(#Rows, #Columns, #Links)`.

This allows nestedness values to be normalised as NODF/max(NODF) following Song et al (2017). To control for connectance and network size, Song et al. (2017) suggest an additional normalisation that can be used: (NODF/max(NODF))/(C * log(S)) where C is the network connectance and S is the geometric mean of the number of plants and pollinators in the network.

`maxnodf` has three algorithms for finding the maximum nestedness of a bipartite network. These can be set using the `quality` argument. Lower quality settings are faster, but find worse optima. Higher quality settings are slower, but find better optima.

* `quality = 0`, uses a greedy algorithm.
* `quality = 1`, uses a greedy algorithm plus hillclimbing.
* `quality = 2`, uses a simulated annealing algorithm, with the greedy algorithm output as the start point. Best results, but requires the most computation time.

## Example
```{r, eval = FALSE}
m <- matrix(0,10,10) # initialise an empty network
m[1,] <- 1 # ensure all row species have at least one link
m[,1] <- 1 # ensure all column species have at least one link
m[2:10,2:10] <- sample(0:1, 9 * 9, replace = TRUE) # randomise the rest of the matrix
maxnodf(web = m, quality = 2) # calculate the maximum nestedness
```

## License
The code is released under the MIT license (see `LICENSE` file).

## References
Song, C., Rohr, R.P. and Saavedra, S., 2017. Why are some plant–pollinator networks more nested than others? Journal of Animal Ecology, 86(6), pp.1417-1424.